---
title: "*Culex pipiens* egg production depends upon vertebrate blood host species"
author: "Cuadera, M.K, and Fritz, M. L."
date: "7/6/2018"
output: github_document
---



####Background####
The *Culex pipiens* species is part of the Culex assemblage. Within the species itself, there exists two major bioforms: form pipiens and form molestus. *Cx. pip* form pipiens are aboveground mosquitoes that feed primarily in avian blood, while the *Cx. pip* form molestus are underground mosquitoes that feed primarily in mammalian blood. Though host preference has been previously demostrated, the reason as to why these preferences exist is unknown. The purpose of this work is to analyze the effects of different mammalian and avian host blood on the fecundity of above ground and below ground *Cx. pipiens* mosquitoes to see whether mosquitoes produce the most eggs when feeding on their preferred hosts.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Analytical Methods

##### Summary function
The function below takes the main data frame and outputs a new dataframe that includes the mean, standard deviation, and number of replicates. 
```{r}
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE),
      No_Replicates = length(x[[col]]))}
  
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}
```
##### Bootstrapping function
```{r}
boot.fn <- function(x, N=5000) {
  Int.1 <- replicate(N, mean(sample(x, size= length(x), replace=T)))
  Int.CI <- quantile(Int.1, probs=c(0.025,0.975))
  Int.CI
}
```

#####Setting up the main dataset
```{r, results = "hide"}
#Mervin uses this
#Exp_1_data <- read.csv(file = "~/Desktop/Mervin-Culex_Fitness_By_Blood/data/Exp1_Host_Blood_Fitness_Study_Replicates.csv",header = T)
#Megan uses this
Exp_1_data <- read.csv(file = "~/Desktop/Culex_Fitness_By_Blood/data/Exp1_Host_Blood_Fitness_Study_Replicates.csv",header = T)
head(Exp_1_data)
Exp_1_data$No_eggs_per_fem <- Exp_1_data$Total_Egg_Produced/Exp_1_data$No_Fed
Fed_only <- subset(Exp_1_data, No_Fed > 0)

str(Fed_only)
Fed_only_mammalian <- subset(Fed_only, Host_Type =="mammalian")
Fed_only_avian <- subset(Fed_only, Host_Type == "avian")
Fed_only_molestus <- subset(Fed_only, Form =="m")
Fed_only_pip <- subset(Fed_only, Form =="p")

Fed_only_mol_bovine <- subset(Fed_only_molestus, Treatment =="Bovine")
Fed_only_mol_rabbit <- subset(Fed_only_molestus, Treatment =="Rabbit")
Fed_only_mol_equine <- subset(Fed_only_molestus, Treatment =="Equine")
Fed_only_mol_goose <- subset(Fed_only_molestus, Treatment =="Goose")
Fed_only_mol_chicken <- subset(Fed_only_molestus, Treatment =="Chix")
Fed_only_mol_turkey <- subset(Fed_only_molestus, Treatment =="Turkey")

Fed_only_pip_bovine <- subset(Fed_only_pip, Treatment =="Bovine")
Fed_only_pip_rabbit <- subset(Fed_only_pip, Treatment =="Rabbit")
Fed_only_pip_equine <- subset(Fed_only_pip, Treatment =="Equine")
Fed_only_pip_goose <- subset(Fed_only_pip, Treatment =="Goose")
Fed_only_pip_chicken <- subset(Fed_only_pip, Treatment =="Chix")
Fed_only_pip_turkey <- subset(Fed_only_pip, Treatment =="Turkey")
```

#####Setting up the libraries needed
```{r echo=T, error = FALSE, results='hide'}
library (lme4)
library(ggplot2)
library(car)
library(plyr)
```

#####Table summary of the data
Table summary with the standard error calculated.
```{r, results= "hide", error=FALSE}
Exp_1_Data_Summary <- data_summary(Fed_only, varname = "No_eggs_per_fem", groupnames = c("Strain", "Treatment"))
Exp_1_Data_Summary$se <- Exp_1_Data_Summary$sd / sqrt(Exp_1_Data_Summary$No_Replicates)
```
```{r}
head(Exp_1_Data_Summary)
```
```{r, results = "hide"}
Exp_1_Data_Host_Type <-data_summary(Fed_only, varname = "No_eggs_per_fem", groupnames = c("Strain", "Host_Type"))
Exp_1_Data_Host_Type$se <- Exp_1_Data_Host_Type$sd / sqrt(Exp_1_Data_Host_Type$No_Replicates)
```
```{r}
head(Exp_1_Data_Host_Type)
```

####Graphs of the Data

#####Average egg production when looking at host type alone
Looking at host type alone, all strains seem to do well when fed avian blood. Error bars represent standard errors (alpha = 0.05)
```{r, echo = FALSE}
Host_Type_bar <-ggplot(Exp_1_Data_Host_Type, aes(x=Strain, y=No_eggs_per_fem, fill=Host_Type)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=No_eggs_per_fem-se, ymax=No_eggs_per_fem+se), width=.2,
                position=position_dodge(.9)) + labs(x="Strain", y = "Average Eggs", fill = "Host Type")+
  theme_classic()+ scale_color_brewer(palette = "Spectral")+theme(text=element_text(family="Calibri"))
print(Host_Type_bar)
```

#####Average egg production per treatment
The graph below summarizes the average egg production for all treatments. Error bars represent standard error (alpha = 0.05)
```{r, echo = FALSE}
whole_data_bar <-ggplot(Exp_1_Data_Summary, aes(x=Strain, y=No_eggs_per_fem, fill=Treatment)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=No_eggs_per_fem-se, ymax=No_eggs_per_fem+se), width=.2,
                position=position_dodge(.9)) + labs(x="Strain", y = "Average Eggs", fill = "Treatment")+
  theme_classic()+ scale_color_brewer(palette = "Spectral")+theme(text=element_text(family="Calibri"))
print(whole_data_bar)
```

####Linear Model With Mixed Effects for the Data
The linear mixed effects model attempts to explain egg production with respect to strain and treatment, including their interaction and the random effect caused by performing the experiment in replicates over a wide span of time. 
```{r, error = FALSE, results = "hide"}
model_Full1 <- lmer(No_eggs_per_fem ~ 1 + Strain*Treatment + (1|Replicate), data = Fed_only)
summary(model_Full1)
```

####Statistical Analysis
#####ANOVA 
A two-way ANOVA indicates that host and strain are signficant factors that impact egg production. The interaction between host and strain is moderately significant. 
```{r}
Host_Anova <- aov(No_eggs_per_fem ~ Host_Type*Strain, data = Fed_only)
summary(Host_Anova)
TukeyHSD(Host_Anova)
```
#####T-Test
Treatments were compared within the *molestus* form. A significant difference was found in the chicken and turkey treatments, with CAL1 having the greatest average egg production for both.
```{r}
t.test(No_eggs_per_fem~Strain, data = Fed_only_mol_bovine)
t.test(No_eggs_per_fem~Strain, data = Fed_only_mol_chicken)
t.test(No_eggs_per_fem~Strain, data = Fed_only_mol_equine)
t.test(No_eggs_per_fem~Strain, data = Fed_only_mol_goose)
t.test(No_eggs_per_fem~Strain, data = Fed_only_mol_rabbit)
t.test(No_eggs_per_fem~Strain, data = Fed_only_mol_turkey)
```
Within the *pipiens* form, a significant difference was found in the bovine, chicken, and equine treatments, with Evanston having the greatest average egg production for those treatments. 
```{r}
t.test(No_eggs_per_fem~Strain, data = Fed_only_pip_bovine)
t.test(No_eggs_per_fem~Strain, data = Fed_only_pip_chicken)
t.test(No_eggs_per_fem~Strain, data = Fed_only_pip_equine)
t.test(No_eggs_per_fem~Strain, data = Fed_only_pip_goose)
t.test(No_eggs_per_fem~Strain, data = Fed_only_pip_rabbit)
t.test(No_eggs_per_fem~Strain, data = Fed_only_pip_turkey)
```

####Conclusion
Across strains, both forms seem to produce the most eggs when feed avian blood.  